Attribute fusion in a latent process model for time series of graphs

### 2011-04-26 ssp2011.tex
Add section on max degree

\subsection{Graphs Invariants}
In this paper we consider the problem of detecting chatter anomalies
in time series of graphs using attributed invariants. The particular
invariants of interests are the size, max degree, scan, and number of
triangles. Specifically, we consider linear attribute fusion with
parameter $\lambda \in \mathbb{R}^{K}_{\geq 0}$ via

@@

  \mathrm{size}_{\lambda}(G(t)) = \sum_{k=1}^{K} \lambda_k \sum_{u,v}

@@

window. The main object of our interests are then the normalized fusion
statistic $T_{\lambda}^{l}(t)$ as defined by

@@

  \caption{Temporal standardization: when testing for change at time $t$, the recent past (graphs $G(t - l), \dots, G(t-1)$ is used to standardize the invariants}

@@

@@

\cite{lee:_laten_proces_model_time_attrib_random_graph} that when
$J_\lambda(t) = \mathrm{size}_{\lambda}(t)$, the normalized fusion
statistic $T_{\lambda}^{l}(t)$ follows a $t$-distribution with $l - 1$

@@

    \xi = (1+1/l)\tbinom{n}{2} \eta_{00} +

@@

As $n \rightarrow \infty$, the random variable

@@

      0 & \text{otherwise}

@@

\begin{theorem}
  \label{thm:1}
  Let $\lambda^{*}$ be the eigenvector corresponding to the largest
  eigenvalue of $ \mathbf{A}^{-1/2} \mathbf{B} \mathbf{A}^{-1/2}$,
  i.e., 
 \begin{equation}
   \label{eq:9}
  \lambda^{*} = \argmax_{\lambda \colon \| \lambda \| = 1}
  \lambda^{T} \mathbf{A}^{-1/2} \mathbf{B} \mathbf{A}^{-1/2}
  \lambda.
 \end{equation}
 We then have
 \begin{gather}
   \label{eq:8}
 \argmax_{ \|\lambda\| = 1}
 \mu_{\lambda} \, \cap \, \{\lambda^{*}, - \lambda^{*}\} \not = \emptyset \\
 \argmin_{ \| \lambda \| = 1} \mu_{\lambda} \, \cap \, \{\lambda^{*}, -
 \lambda^{*}\} \not = \emptyset.
 \end{gather}
\end{theorem}
\bibliography{ssp2011}
\end{document}

>>>>

Let $\mathscr{V}$ be a random dot process model with change parameters
$(t^{*}, m, \pi_0, \pi_1, \mathbf{Q}_0, \mathbf{Q}_1)$. The
change-parameters encapsulate a notion of chatter anomalies, i.e., a
subset of vertices of $\mathscr{V}$ with altered communication
behaviour in an otherwise stationary setting as depicted in Fig.~\ref{fig:notional_change_point}. 
  \includegraphics[width=8cm]{graphics/Fig1-SSP2011.pdf}
  \caption{Notional depiction of the problem of change-point detection in a time-series of graphs}
  \label{fig:notional_change_point}
\\ \noindent
We are interested in
the problem of testing, for a $t \in \mathbb{N}$, the hypotheses that
$t$ is the change-point of $\mathscr{V}$, namely
  \mathscr{H}_0 \colon t^{*} > t \\
  \mathscr{H}_A \colon t^{*} = t
%\subsection{Graphs Invariants}
%In this paper we consider the problem of detecting chatter anomalies
This will be done using the notion of fusion of attributed graph
invariants. The particular invariants of interests are the size
$\mathcal{E}$, number of triangles $\tau_{\lambda}$, scan $\Psi$, and
max degree $\Delta$. Specifically, we consider linear attribute
fusion with parameter $\lambda \in \mathbb{R}^{K}$ via

@@

  \mathcal{E}_{\lambda}(G(t)) = \sum_{k=1}^{K} \lambda_k \sum_{u,v}

@@

window. Our main interests is in the normalized fusion
statistic $T_{\lambda}^{l}(t)$ as depicted in Fig.~\ref{fig:temporal}, namely

@@

  \caption{Temporal standardization: when testing for change at time $t$, the recent past (graphs $G(t - l), \dots, G(t-1))$ is used to standardize the invariants}

@@

\subsection{Power estimates for $\mathcal{E}_\lambda(t)$}

@@

\cite{lee:_laten_proces_model_time_attrib_random_graph} that for the
case where $J_\lambda(t)$ is $\mathcal{E}_{\lambda}(t)$, $T_{\lambda}^{l}(t)$ 
follows a $t$-distribution with $l - 1$

@@

    \xi = \tfrac{l+1}{l}\tbinom{n}{2} \eta_{00} +

@@

  Define the random variable $\psi^{l}_{\lambda}(t)$ by
    \psi^{l}_{\lambda}(t) = \begin{cases}
      \sqrt{\frac{\langle \lambda, \tbinom{n}{2} \eta_{00}
            \lambda \rangle}{\langle \lambda, \xi
            \lambda \rangle}} T_{\lambda}^{l}(t) & \text{if $t = t^{*}$} \\
      \sqrt{\frac{l}{l + 1}} T_{\lambda}^{l}(t)& \text{if $t < t^{*}$}
As $n \rightarrow \infty$,

@@

      0 & \text{if $t < t^{*}$}

@@

We are interested in finding the $\lambda$ that will maximize the
power of the test. The power approximation is determined by various
factors but the dominating factor is likely to be
$\mu_{\lambda}$. Theorem \ref{thm:1} extends a result in
\cite{lee:_laten_proces_model_time_attrib_random_graph} for $K = 2$ to the case
where $K \geq 2$. 
\begin{theorem}
  \label{thm:1}
  Let $\zeta$ and $\xi$ be as defined in Lemma \ref{lem:1}.  Suppose
  that $\xi$ is also positive definite. Let $\nu^{*}$ be the
  normalized eigenvector corresponding to the largest eigenvalue of $ \xi^{-1/2}
  \zeta \zeta^{T} \xi^{-1/2}$, i.e.,
   \label{eq:9}
  \nu^{*} = \argmax_{\nu \colon \| \nu \| = 1}
  \nu^{T} \xi^{-1/2} \zeta \zeta^{T} \xi^{-1/2}
  \nu.
 Then $\lambda^{*} = \tfrac{\xi^{-1/2} \nu^{*}}{\|\xi^{-1/2} \nu^{*} \|}$ satisfies
   \label{eq:8}
 \argmax_{ \|\lambda\| = 1}
 \mu_{\lambda} \, \cap \, \{\lambda^{*}, - \lambda^{*}\} \not = \emptyset \\
 \argmin_{ \| \lambda \| = 1} \mu_{\lambda} \, \cap \, \{\lambda^{*}, -
 \lambda^{*}\} \not = \emptyset.
\end{theorem}
\subsection{Power estimates for $\Delta_{\lambda}(t)$}
The analysis of the power estimates for $\Delta_{\lambda}(t)$ depends
on some preliminary results which we now state. We denote by
$\mathcal{G}(\alpha, \beta)$ the Gumbel distribution with location
parameter $\alpha$ and scale parameter $\beta$.  
\begin{proposition}
  \label{prop:3}
  Let $a_n$ and $b_n$ be constants given by
  \begin{align*}
    a_n &= (2 \log{n})^{1/2}\Bigl(1 - \frac{\log{\log{n}}}{4 \log{n}} -
    \frac{\log{2\pi^{1/2}}}{2 \log{n}} \Bigr) \\ 
    b_n &= (2 \log{n})^{-1/2}
  \end{align*}
  Then as $n \rightarrow \infty$ and $m = \Omega( \sqrt{n \log{n}})$,
  we have
    \Delta_{\lambda}(t) \rightsquigarrow \begin{cases}
      \mathcal{G}(a_n \sigma_1 + \mu_1, b_n \sigma_1 ) & \text{for $t < t^{*}$.} \\
     \mathcal{G}(a_n \sigma_2 + \mu_2, b_n \sigma_2) &
     \text{for $t = t^{*}$.}
  where
  \begin{align*}
    \mu_1 &= (n-1)\langle x, \pi_{00} \rangle \\
    \sigma_1 &= \sqrt{(n-1)\langle x, \eta_{00} x \rangle} \\
    \mu_2 &= \,\, (m - 1) \langle x, \pi_{11} \rangle + (n-
    m)\langle x, \pi_{01} \rangle
    \\ \sigma_2 &= \sqrt{ (m - 1) \langle x, (\eta_{11} x \rangle + (n -
      m) \left<x, \eta_{01} x\right>}
    \end{align*}
\end{proposition}
  \label{lem:2}
  \mathbb{R}^{K}$. For sufficiently large $n$ and sufficiently large
  $l$, the variable $T_{\lambda}^{l}(t)$ has approximately a Gumbel
  distribution with location $\rho_{\lambda}$ and scale $\varsigma_{l}$ where
  \label{eq:8}
  \rho_{\lambda} = \begin{cases}
    \frac{ (a_n + b_n \gamma)(\sigma_2 - \sigma_1) + (\mu_2 -
    \mu_1)}{\tfrac{\pi}{\sqrt{6}}\sqrt{(1 + 1/m)\sigma_1^2 +
      (\sigma_2^2 - \sigma_1^2)}} & \text{if $t = t_1$} \\ 
  0 & \text{if $t < t_1$}
$\gamma \approx 0.57721$ is the Euler-Mascheroni constant, and
\appendices
\section{Proofs of some stated results}
\begin{proof}[Theorem \ref{thm:1}]
  The maximizer of $\mu_\lambda$ also maximizes
  \begin{equation*}
    \mu_{\lambda}^{2} = \frac{\lambda^{T} \zeta \zeta^{T}
      \lambda}{\lambda^{T} \xi \lambda}
  \end{equation*}
  Because $\xi$ is positive definite, there exist a positive definite matrix
  $\xi^{1/2}$ such that $\xi^{1/2} \xi^{1/2} = \xi$. Letting $\nu = \xi^{1/2}
  \lambda$, the above expression can be rewritten as
  \begin{equation*}
    \mu_{\lambda}^{2} = \frac{\nu^{T} \xi^{-1/2} \zeta \zeta^{T}
      \xi^{-1/2} \nu}{ \nu^{T} \nu}
  \end{equation*}
  The claim then follows directly from the Rayleigh-Ritz theorem for
  Hermitean matrices.
\end{proof}
\bibliography{ssp2011}
\end{document}
